High-level cognition is supported by at least second order dynamic correlations in neural activity patterns

### 2019-09-02 paper/boneyard.tex
working on fmri temporal decoding results

>>>>

We next evaluated if our model of high-order correlations in brain activity can capture cognitively relevant brain patterns. We performed a decoding analysis, using cross validation to estimate (using other participants’ data) which parts of the story each weighted-mixture of higher-order brain activity pattern corresponded to (see \textit{Materials and methods}). We note that our primary goal was not to achieve perfect decoding accuracy, but rather to use decoding accuracy as a benchmark for assessing whether different neural features specifically capture cognitively relevant brain patterns.

Separately for each experimental condition, we divided participants
into two groups. For the zeroth order, we computed the mean factor
activity for each group.  For all subsequent orders up to the tenth
order, we computed the mean approximated dynamic ISFC of factor
activity for each group (see \textit{Materials and methods}), and
combined in a weighted mixutre with all previous orders
(i.e. cross-validation for the second
order contained a weighted-mixture of zeroth, first, and second order
(Fig.~\ref{fig:decoding_level}, c.\&f.).  For
each order, we correlated the group 1 activity patterns with group 2
activity patterns.  We then subdivided the group 1 to obtain an
optimal weighting parameter for each order’s correlation matrix using
the same cross validation method. We used the optimal weighting
parameters to obtain a weighted-mixture (see \textit{Materials and methods}) of each order’s correlation
matrix. Using these correlations, we labeled the group 1 timepoints
using the group 2 timepoints with which they were most highly
correlated; we then computed the proportion of correctly labeled group
1 timepoints. (We also performed the symmetric analysis whereby we
labeled the group 2 timepoints using the group 1 timepoints as a
template.) We repeated this procedure 100 times (randomly re-assigning
participants to the two groups each time) to obtain a distribution of
decoding accuracies for each experimental condition. There were 272
timepoints for paragraph condition, 300 timepoints for intact and word
conditions, and 400 timepoints for rest condition,  so chance
performance on this decoding test is was $\frac{1}{272}$,
$\frac{1}{300}$, and $\frac{1}{400}$ respectively.

We repeated this process for each set of parameters, varying kernel
type and width, and averaged over the reduction technique used to
approximate the higher-order correlations (PCA Fig.~\ref{fig:decoding_level},  a.-c. and eigenvector
centrality Fig.~\ref{fig:decoding_level},  d.-f.).  Since there is no ground truth in these analyses, and we did not know
which parameters best capture the data, we instead report a robustness
search by averaging over the parameters and reporting which results
consistently showed up across all parameters.
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/brain_plots.pdf}
  \caption{\textbf{Average correlations by order for the intact listening condition.} Using eigenvector
    centrality to approximate higher-order correlations for the
    intact, paragraph scrambled, word scrambled, and rest condition.
    We plot the
    strongest 50\% absolute value
    mean correlation for \textbf{a.-d. first order, e.-h. second order,
      i.-l. third order, and m.-p. fourth order }, representing the degree of agreement by
    location pair over time.  To demonstrate how this method is
    computationally scalable, we also approximated
    \textbf{a.-d. fifteenth order} dynamic correlation, which would be
    possible to compute using conventional methods since it would require more bits to represent the solution than there are molecules in the universe.}
  \label{fig:brain_plots}
\end{figure}

The two methods used to approximate the higher-order correlations (PCA Fig.~\ref{fig:decoding_level},  a.-c. and eigenvector
centrality Fig.~\ref{fig:decoding_level},  d.-f.) capture different
facets of the activity patterns.  Using PCA, the higher-order
correlations are all linked to the original activity patterns, whereas
eigenvectory centrality breaks the immediate link with specific brain
areas and instead characterizes the position of the nodes in the
network that are similar over time.

We found for both PCA and eigenvector centrality, during the intact
condition in the experiment, classifiers that incorporated
higher-order correlations yielded consistently higher accuracy than
classifiers trained only on lower-order patterns
(Fig.~\ref{fig:decoding_level}, a.\&d.).  We plot the average
correlations for up to the fourth order for the intact condition
(Fig.~\ref{fig:brain_plots}) representing the degree of agreement by
location pair over time.  By contrast, we found that incorporating
higher-order (greater than first order) correlations did not further
improve decoding accuracy for the other listening conditions or rest
condition.  This suggests that the cognitive processing that supported
the most cognitively rich condition involved higher-order network
dynamics.

%synthetic data: we can recover dynamic correlations using different
%kernels
% wider kernels (Laplace, Gaussian) performed best when the
% correlations varied in a temporally structured way
% delta function kernels were best when the correlations varied
% randomly over time
% overall we recovered the structure we embedded in the data under all
% synthetic datasets we tested.  our approach performed best when the
% underlying correlational structure was either constant or gradually
% changing.  our approach performed worst when the underlying
% correlations varied randomly for one timepoint to the next.

\subsubsection*{Synthetic data}

To assess the performance of dynamic correlation recovery using
timecorr, we varied width the kernel and the specific structure of the
data. We applied timecorr, using delta and gaussian kernels
Fig.~\ref{fig:kernels}) to each of the following  
synthetic datasets: constant, random, ramping, and block.  We then correlated each recovered
correlation matrix with the ground truth.

For the constant synthetic dataset, a gaussian kernel (width=10)
outperformed the delta kernel (Fig.~\ref{fig:synthetic_data},  a.).  This is in contrast with the random
synthetic dataset, for which the delta kernel best captures the
rapidly changing structure (Fig.~\ref{fig:synthetic_data},  b.). For
the ramping synthetic dataset, the slow changing strucutre within the
data is best
captured by the gaussian kernel and the best recovery occurs in the
middle (Ramping, Fig.~\ref{fig:synthetic_data},
c.). In addition to comparing the timecorr recovered correlation
matrices to the ground truth, we
further compared the ramping recovered correlation matrices to only the first random covariance matrix $K_{1}$
(First, Fig.~\ref{fig:synthetic_data},  c.) and to only the last
random covariance matrix $K_{2}$ (Last, Fig.~\ref{fig:synthetic_data},
c.), both of which perform best at the beginning and end respectively.

Similary for the block sythetic dataset, we compared the timecorr
recovered correlation matrices to the ground truth as well as to each
block-specific covariance matrix (Block 1-5,
Fig.~\ref{fig:synthetic_data},  d.).  Although the structure is
changing by block, the gaussian kernel once again outperforms the
delta kernel.  Performance does however drop near even boundaries for
when using the gaussian kernel.

### 2019-09-02 paper/main.tex
working on fmri temporal decoding results

Another central question pertains to the extent to which the neural
code is carried by activity patterns that directly reflect ongoing
cognition~\citep[e.g., following][]{HaxbEtal01, NormEtal06}, versus
the dynamic properties of the network structure itself, independent of
specific activity patterns in any given set of regions~\citep[e.g.,
following][]{BassEtal06}.  For example, graph theoretic measures such
as centrality and degree~\citep{BullSpor09} may be used to estimate
how a given brain structure is ``communicating'' with other
structures, independently of the specific neural representations
carried by those structures.  If one considers a brain region's graph
theoretic position in the network (e.g., its eigenvector centrality)
as a dynamic property, one can compare how the positions of different
regions are correlated, and/or how those patterns of correlations change
over time.  We can also compute higher-order patterns in these
correlations to characterize homologous subgraphs in the connectome
that display similar changes in their constituent brain structures'
interactions with the rest of the brain.

@@

Given a $T$ by $\frac{K^2 - K}{2}$ matrix of dynamic correlations,
$\mathbf{Y}_n$, we propose two general approaches to computing a $T$
by $K$ low-dimensional representation of these correlations,
$\mathbf{X}_n$.  The first approach uses dimensionality reduction
algorithms to project $\mathbf{Y}_n$ onto a $K$-dimensional space.
The second approach uses graph-theoretic measures to characterize the
relative positions of each feature ($k \in \left[1, ..., K \right]$)
in the network defined by the correlation matrix at each timepoint.

@@

\subsubsection*{Graph theoretic approaches to computing
  $\mathbf{X}_n$}
The above dimensionality reduction approaches to approximating a given
$\mathbf{Y}_n$ with a lower-dimensional $\mathbf{X}_n$ preserve a
(potentially recombined and transformed) mapping back to the original
data in $\mathbf{Y}_0$.  We also explore graph theoretic approaches
that forgo a preserved mapping back to the original data in favor of
preserving each feature's relative \textit{position} in the broader
network of interactions and connections.  To illustrate the
distinction between the two general approaches we explore, suppose a
network comprises nodes $A$, $B$, and $C$.  If $A$ and $B$ exhibit
uncorrelated activity patterns, the functional connection between
them will be (by definition) close to 0.  However, if $A$ and $B$ each
interact with $C$ in similar ways, we might attempt to capture those
similarities using a measure that reflects the how $A$ and $B$
interact in the network.

In general, graph theoretic measures take as input a matrix of
interactions (e.g., using the above notation, an $K$ by $K$
correlation matrix or binarized correlation matrix reconstituded from
a single timepoint's row of $\mathbf{Y}$) and return as output a set
of $K$ measures describing how each node (feature) sits within that
correlation matrix with respect to the rest of the population.  Widely
used measures include betweeness centrality~\citep[the proportion of
shortest paths between each pair of nodes in the population that
involves the given node in question; e.g., ][]{Newm05, OpsaEtal10,
  Bart04, GeisEtal08, Free77}; diversity and
dissimilarity~\citep[characterizations of how differently connected a
given node is from others in the population; e.g., ][]{Rao82, Lin09,
  RicoSzei06}; Eigenvector centrality and pagerank
centrality~\citep[measures of how influential a given node is within
the broader network; e.g., ][]{Newm08, Bona07, LohmEtal10,
  HaluEtal13}; transfer entropy and flow coefficients~\citep[a measure
of how much information is flowing from a given node to other nodes in
the network; e.g., ][]{HoneEtal07, Schr00}; $k$-coreness
centrality~\citep[a measure of the connectivity of a node within its
local sub-graph; e.g., ][]{AlvaEtal05, ChriFowl10}; within-module
degree~\citep[a measure of how many connections a node has to its
close neighbors in the network; e.g., ][]{RubiSpor10}; participation
coefficient~\citep[a measure of the diversity of a node's connections
to different sub-graphs in the network; e.g., ][]{RubiSpor10}; and
sub-graph centrality~\citep[a measure of a node's participation in all
of the network's sub-graphs; e.g., ][]{EstrRodr05}; among others.

For a given graph theoretic measure,
$\eta: \mathbb{R}^{K \times K} \rightarrow \mathbb{R}^K$, we can use
$\eta$ to tranform each row of $\mathbf{Y}_n$ in a way that
characterizes the corresponding graph-theoretic properties of each
column.  This results in a new $T$ by $K$ matrix, $\mathbf{X}_n$, that
reflects how the features reflected in the columns of $\mathbf{Y}_n$
participate in the network during each timepoint (row).

@@

\subsection*{Recovering known dynamic correlations from synthetic data}

%synthetic data: we can recover dynamic correlations using different
%kernels
% wider kernels (Laplace, Gaussian) performed best when the
% correlations varied in a temporally structured way
% delta function kernels were best when the correlations varied
% randomly over time
% overall we recovered the structure we embedded in the data under all
% synthetic datasets we tested.  our approach performed best when the
% underlying correlational structure was either constant or gradually
% changing.  our approach performed worst when the underlying
% correlations varied randomly for one timepoint to the next.

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/synthetic_data.pdf}
  \caption{\textbf{Recovering known dynamic correlations from
      synthetic data.}  Each panel displays the average correlations between
    the vectorized upper triangles of the recovered correlation matrix
    at each timepoint and either the true underlying correlation at each timepoint or a
    reference correlation matrix.  (The averages are taken across 10
    different randomly generated synthetic datasets.) Error ribbons
    denote 95\% confidence intervals (taken across datasets). Different colors denote different
    kernel shapes, whereas the shading within each color type denotes
    kernel widths.  For a complete description of each synthetic
    dataset, see \textit{Synthetic data}.  \textbf{a. Constant
      correlations.} These datasets have a stable (unchanging)
    underlying correlation matrix.  \textbf{b. Random
      correlations.}  These datasets are generated using a new
    independently drawn correlation matrix at each new timepoint.  \textbf{c. Ramping
      correlations.}  These datasets are generated by smoothly varying
    the underlying correlations between the randomly drawn
    correlation matrices at first and last timepoints.  The left
    panel displays the correlations between the recovered dynamic
    correlations and the underlying ground truth correlations.  The
    middle panel compares the recovered correlations with the
    \textit{first} timepoint's correlation matrix.  The right panel
    compares the recovered correlations with the
    \textit{last} timepoint's correlation matrix.
    \textbf{d. Event-based correlations.}  These datasets are each
    generated using five randomly drawn correlation matrices that each
    remain stable for a fifth of the total timecourse.  The left panel
    displays the correlations between the recovered dynamic
    correlations and the underlying ground truth correlations.  The
    right panels compare the recovered correlations with the
    correlation matrices unique to each event.}
  \label{fig:synthetic_data}
\end{figure}

@@

\subsubsection*{Synthetic data}

To assess the performance of dynamic correlation recovery using
timecorr, we varied width the kernel and the specific structure of the
data. We applied timecorr, using delta and gaussian kernels
Fig.~\ref{fig:kernels}) to each of the following  
synthetic datasets: constant, random, ramping, and block.  We then correlated each recovered
correlation matrix with the ground truth.

For the constant synthetic dataset, a gaussian kernel (width=10)
outperformed the delta kernel (Fig.~\ref{fig:synthetic_data},  a.).  This is in contrast with the random
synthetic dataset, for which the delta kernel best captures the
rapidly changing structure (Fig.~\ref{fig:synthetic_data},  b.). For
the ramping synthetic dataset, the slow changing strucutre within the
data is best
captured by the gaussian kernel and the best recovery occurs in the
middle (Ramping, Fig.~\ref{fig:synthetic_data},
c.). In addition to comparing the timecorr recovered correlation
matrices to the ground truth, we
further compared the ramping recovered correlation matrices to only the first random covariance matrix $K_{1}$
(First, Fig.~\ref{fig:synthetic_data},  c.) and to only the last
random covariance matrix $K_{2}$ (Last, Fig.~\ref{fig:synthetic_data},
c.), both of which perform best at the beginning and end respectively.

Similary for the block sythetic dataset, we compared the timecorr
recovered correlation matrices to the ground truth as well as to each
block-specific covariance matrix (Block 1-5,
Fig.~\ref{fig:synthetic_data},  d.).  Although the structure is
changing by block, the gaussian kernel once again outperforms the
delta kernel.  Performance does however drop near even boundaries for
when using the gaussian kernel.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/decode_level.pdf}
  \caption{\textbf{Decoding by order.
      a.\&d. Relative decoding accuracy by order.} Ribbons of each
    color display cross-validated decoding performance averaged over
    all parameters for each condition (intact, paragraph, word,
    and rest) using PCA (a.) or eigenvector centrality (d.) to
    approximate correlations. Decoders were trained using increasingly more
    higher-order information and this ribbons are displayed relative
    to chance at 0. The red dots indicates
    maximum decoding accuracy for each condition. \textbf{b.\&e. Z-transformed decoding accuracy
      by order.}  We Z-transformed the decoding accuracy by order to better
    visualize the order with the maximum decoding accuracy for each
    condition using PCA (b.) or eigenvector centrality (e.) to
    approximate correlations. \textbf{c.\&f. Optimized weights.} Bar heights indicate the
    optimized mixing paramete $\phi$ of each contributing order up to
    and including the order with the maximum
    decoding accuracy for each contributing order using PCA (c.) or eigenvector centrality (f.) to
    approximate correlations.
  For the order with maximum decoding accuracy by condition, we show
  barplots of the optimized weights $\phi$  for each contributing
  order.}
  \label{fig:decoding_level}
\end{figure}

We next evaluated if our model of high-order correlations in brain activity can capture cognitively relevant brain patterns. We performed a decoding analysis, using cross validation to estimate (using other participants’ data) which parts of the story each weighted-mixture of higher-order brain activity pattern corresponded to (see \textit{Materials and methods}). We note that our primary goal was not to achieve perfect decoding accuracy, but rather to use decoding accuracy as a benchmark for assessing whether different neural features specifically capture cognitively relevant brain patterns.

Separately for each experimental condition, we divided participants
into two groups. For the zeroth order, we computed the mean factor
activity for each group.  For all subsequent orders up to the tenth
order, we computed the mean approximated dynamic ISFC of factor
activity for each group (see \textit{Materials and methods}), and
combined in a weighted mixutre with all previous orders
(i.e. cross-validation for the second
order contained a weighted-mixture of zeroth, first, and second order
(Fig.~\ref{fig:decoding_level}, c.\&f.).  For
each order, we correlated the group 1 activity patterns with group 2
activity patterns.  We then subdivided the group 1 to obtain an
optimal weighting parameter for each order’s correlation matrix using
the same cross validation method. We used the optimal weighting
parameters to obtain a weighted-mixture (see \textit{Materials and methods}) of each order’s correlation
matrix. Using these correlations, we labeled the group 1 timepoints
using the group 2 timepoints with which they were most highly
correlated; we then computed the proportion of correctly labeled group
1 timepoints. (We also performed the symmetric analysis whereby we
labeled the group 2 timepoints using the group 1 timepoints as a
template.) We repeated this procedure 100 times (randomly re-assigning
participants to the two groups each time) to obtain a distribution of
decoding accuracies for each experimental condition. There were 272
timepoints for paragraph condition, 300 timepoints for intact and word
conditions, and 400 timepoints for rest condition,  so chance
performance on this decoding test is was $\frac{1}{272}$,
$\frac{1}{300}$, and $\frac{1}{400}$ respectively.

We repeated this process for each set of parameters, varying kernel
type and width, and averaged over the reduction technique used to
approximate the higher-order correlations (PCA Fig.~\ref{fig:decoding_level},  a.-c. and eigenvector
centrality Fig.~\ref{fig:decoding_level},  d.-f.).  Since there is no ground truth in these analyses, and we did not know
which parameters best capture the data, we instead report a robustness
search by averaging over the parameters and reporting which results
consistently showed up across all parameters.
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/brain_plots.pdf}
  \caption{\textbf{Average correlations by order for the intact listening condition.} Using eigenvector
    centrality to approximate higher-order correlations for the
    intact, paragraph scrambled, word scrambled, and rest condition.
    We plot the
    strongest 50\% absolute value
    mean correlation for \textbf{a.-d. first order, e.-h. second order,
      i.-l. third order, and m.-p. fourth order }, representing the degree of agreement by
    location pair over time.  To demonstrate how this method is
    computationally scalable, we also approximated
    \textbf{a.-d. fifteenth order} dynamic correlation, which would be
    possible to compute using conventional methods since it would require more bits to represent the solution than there are molecules in the universe.}
  \label{fig:brain_plots}
\end{figure}

The two methods used to approximate the higher-order correlations (PCA Fig.~\ref{fig:decoding_level},  a.-c. and eigenvector
centrality Fig.~\ref{fig:decoding_level},  d.-f.) capture different
facets of the activity patterns.  Using PCA, the higher-order
correlations are all linked to the original activity patterns, whereas
eigenvectory centrality breaks the immediate link with specific brain
areas and instead characterizes the position of the nodes in the
network that are similar over time.

We found for both PCA and eigenvector centrality, during the intact
condition in the experiment, classifiers that incorporated
higher-order correlations yielded consistently higher accuracy than
classifiers trained only on lower-order patterns
(Fig.~\ref{fig:decoding_level}, a.\&d.).  We plot the average
correlations for up to the fourth order for the intact condition
(Fig.~\ref{fig:brain_plots}) representing the degree of agreement by
location pair over time.  By contrast, we found that incorporating
higher-order (greater than first order) correlations did not further
improve decoding accuracy for the other listening conditions or rest
condition.  This suggests that the cognitive processing that supported
the most cognitively rich condition involved higher-order network
dynamics.

>>>>

Another central question pertains to the extent to which the neural
code is carried by activity patterns that directly reflect ongoing
cognition~\citep[e.g., following][]{HaxbEtal01, NormEtal06}, versus
the dynamic properties of the network structure itself, independent of
specific activity patterns in any given set of regions~\citep[e.g.,
following][]{BassEtal06}.  For example, graph measures such as
centrality and degree~\citep{BullSpor09} may be used to estimate how a
given brain structure is ``communicating'' with other structures,
independently of the specific neural representations carried by those
structures.  If one considers a brain region's position in the network
(e.g., its eigenvector centrality) as a dynamic property, one can
compare how the positions of different regions are correlated, and/or
how those patterns of correlations change over time.  We can also
compute higher-order patterns in these correlations to characterize
homologous subgraphs in the connectome that display similar changes in
their constituent brain structures' interactions with the rest of the
brain.

@@

Given a $T$ by $\frac{K^2 - K}{2}$ matrix of dynamic correlations,
$\mathbf{Y}_n$, we propose two general approaches to computing a $T$
by $K$ low-dimensional representation of these correlations,
$\mathbf{X}_n$.  The first approach uses dimensionality reduction
algorithms to project $\mathbf{Y}_n$ onto a $K$-dimensional space.
The second approach uses graph measures to characterize the
relative positions of each feature ($k \in \left[1, ..., K \right]$)
in the network defined by the correlation matrix at each timepoint.

@@

\subsubsection*{Graph measure approaches to computing
  $\mathbf{X}_n$}
The above dimensionality reduction approaches to approximating a given
$\mathbf{Y}_n$ with a lower-dimensional $\mathbf{X}_n$ preserve a
(potentially recombined and transformed) mapping back to the original
data in $\mathbf{Y}_0$.  We also explore graph measure approaches
that forgo a preserved mapping back to the original data in favor of
preserving each feature's relative \textit{position} in the broader
network of interactions and connections.  To illustrate the
distinction between the two general approaches we explore, suppose a
network comprises nodes $A$, $B$, and $C$.  If $A$ and $B$ exhibit
uncorrelated activity patterns, the functional connection between
them will be (by definition) close to 0.  However, if $A$ and $B$ each
interact with $C$ in similar ways, we might attempt to capture those
similarities using a measure that reflects the how $A$ and $B$
interact in the network.

In general, graph measures take as input a matrix of
interactions (e.g., using the above notation, an $K$ by $K$
correlation matrix or binarized correlation matrix reconstituded from
a single timepoint's row of $\mathbf{Y}$) and return as output a set
of $K$ measures describing how each node (feature) sits within that
correlation matrix with respect to the rest of the population.  Widely
used measures include betweeness centrality~\citep[the proportion of
shortest paths between each pair of nodes in the population that
involves the given node in question; e.g., ][]{Newm05, OpsaEtal10,
  Bart04, GeisEtal08, Free77}; diversity and
dissimilarity~\citep[characterizations of how differently connected a
given node is from others in the population; e.g., ][]{Rao82, Lin09,
  RicoSzei06}; Eigenvector centrality and pagerank
centrality~\citep[measures of how influential a given node is within
the broader network; e.g., ][]{Newm08, Bona07, LohmEtal10,
  HaluEtal13}; transfer entropy and flow coefficients~\citep[a measure
of how much information is flowing from a given node to other nodes in
the network; e.g., ][]{HoneEtal07, Schr00}; $k$-coreness
centrality~\citep[a measure of the connectivity of a node within its
local sub-graph; e.g., ][]{AlvaEtal05, ChriFowl10}; within-module
degree~\citep[a measure of how many connections a node has to its
close neighbors in the network; e.g., ][]{RubiSpor10}; participation
coefficient~\citep[a measure of the diversity of a node's connections
to different sub-graphs in the network; e.g., ][]{RubiSpor10}; and
sub-graph centrality~\citep[a measure of a node's participation in all
of the network's sub-graphs; e.g., ][]{EstrRodr05}; among others.

For a given graph measure,
$\eta: \mathbb{R}^{K \times K} \rightarrow \mathbb{R}^K$, we can use
$\eta$ to tranform each row of $\mathbf{Y}_n$ in a way that
characterizes the corresponding graph properties of each
column.  This results in a new $T$ by $K$ matrix, $\mathbf{X}_n$, that
reflects how the features reflected in the columns of $\mathbf{Y}_n$
participate in the network during each timepoint (row).

@@

\subsection*{Recovering known dynamic correlations from synthetic data}
We generated synthetic datasets that differed in how the underlying
correlations changed over time.  For each dataset, we applied
Equation~\ref{eqn:timecorr} with a variety of kernel shapes and
widths.  We assessed how well the true underlying correlations at each
timepoint matched the recovered correlations
(Fig.~\ref{fig:synthetic}).  For every kernel and dataset we tested,
our approach recovered the correlation dynamics we embedded into the
data.  However, the quality of these recoveries varied across
different synthetic datasets in a kernel-dependent way.

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/synthetic_data.pdf}
  \caption{\textbf{Recovering known dynamic correlations from
      synthetic data.}  Each panel displays the average correlations
    between the vectorized upper triangles of the recovered
    correlation matrix at each timepoint and either the true
    underlying correlation at each timepoint or a reference
    correlation matrix.  (The averages are taken across 10 different
    randomly generated synthetic datasets of the given category.)
    Error ribbons denote 95\% confidence intervals (taken across
    datasets). Different colors denote different kernel shapes,
    whereas the shading within each color type denotes kernel widths.
    For a complete description of each synthetic dataset, see
    \textit{Synthetic data}.  \textbf{a. Constant correlations.} These
    datasets have a stable (unchanging) underlying correlation matrix.
    \textbf{b. Random correlations.}  These datasets are generated
    using a new independently drawn correlation matrix at each new
    timepoint.  \textbf{c. Ramping correlations.}  These datasets are
    generated by smoothly varying the underlying correlations between
    the randomly drawn correlation matrices at first and last
    timepoints.  The left panel displays the correlations between the
    recovered dynamic correlations and the underlying ground truth
    correlations.  The middle panel compares the recovered
    correlations with the \textit{first} timepoint's correlation
    matrix.  The right panel compares the recovered correlations with
    the \textit{last} timepoint's correlation matrix.
    \textbf{d. Event-based correlations.}  These datasets are each
    generated using five randomly drawn correlation matrices that each
    remain stable for a fifth of the total timecourse.  The left panel
    displays the correlations between the recovered dynamic
    correlations and the underlying ground truth correlations.  The
    right panels compare the recovered correlations with the
    correlation matrices unique to each event.}
  \label{fig:synthetic}
\end{figure}

In general, wide monotonic kernels (Laplace, Gaussian) performed best
when the correlations varied gradually from moment-to-moment
(Figs.~\ref{fig:synthetic}a, c, and d).  \textbf{TODO: Say something
  about kernel widths within a shape.}  In the extreme, as the rate of
change in correlations approaches 0 (Fig.~\ref{fig:synthetic}a), an
infinitely wide kernel would exactly recover the Pearson's correlation
(e.g., compare Eqns.~\ref{eqn:corr} and \ref{eqn:timecorr}).

When the correlation dynamics were unstructured in time
(Fig.~\ref{fig:synthetic}b), a Dirac $\delta$ kernel (infinitely
narrow) performed best.  This is because, when every timepoint's
correlations are independent of the correlations in every other
timepoint, averaging data over time dilutes the available signal.
\textbf{TODO: Say something about kernel widths within a shape.}

\subsection*{Cognitively relevant dynamic high-order correlations in
  fMRI data}
We used across-participant temporal decoders to identify cognitively
relevant neural patterns in fMRI data (see \textit{Forward inference
  and decoding accuracy}).  The dataset we examined~\citep[collected
by][]{SimoEtal16} comprised four experimental conditions that exposed
participants to stimuli that varied systematically in how cognitively
engaging they were.  The \textit{intact} experimental condition had
participants listen to an audio recording of a 10-minute story.  The
\textit{paragraph}-scrambled experimental condition had participants
listen to a temporally scrambled version of the story, where the
paragraphs occurred out of order (but where the same total set of
paragraphs were presented over the full listening interval).  All
participants in this condition experienced the scrambled paragraphs in
the same order.  The \textit{word}-scrambled experimental condition
had participants listen to a temporally scrambled version of the story
where the words in the story occurred in a random order.  All
participants in the word conditions experienced the scrambled words in
the same order.  Finally, in a \textit{rest} experimental condition
participants lay in the scanner with no overt stimulus, with their
eyes open (blinking as needed).  This dataset provided a convenient
means of testing our hypothesis that different levels of cognitive
engagement might be supported by different orders of complex brain
activity dynamics.

In brief, we computed timeseries of dynamic high-order correlations
that were similar across participants in each of two randomly assigned
groups: a training group and a test group.  We then trained
classifiers on the training group's data to match each sample from the
test group with a stimulus timepoint.  Each classifier comprised a
weighted blend of neural patterns that reflected up to
$n^\mathrm{th}$-order dynamic correlations (see \textit{Feature
  weighting and testing}).  We repeated this process for
$n \in \left\{ 0, 1, 2, ..., 10 \right\}$.  Our examinations of synthetic
data suggested that none of the kernels we examined were ``universal''
in the sense of optimally recovering underlying correlations
regardless of the structure of those correlations.  In our analyses of
neural data, we therefore averaged our decoding results over a variety
of kernel shapes and widths in order to identify results that were
robust to specific kernel parameters (also see \textit{Identifying
  robust decoding results}).

Our approach to estimating dynamic high-order correlations requires
mapping the high-dimensional feature space of correlations (a $T$ by
$\mathcal{O}(K^2)$ matrix) onto a lower-dimensional $T$ by $K$ matrix.
We carried out two sets of analyses that differed in how this mapping
was computed.  The first set of analyses used PCA to find a
low-dimensional embedding of the original dynamic correlation matrices
(Fig.~\ref{fig:decoding}a,b).  The second set of analyses
characterized correlations in dynamics of each feature's eigenvector
centrality, but did not preserve the underlying activity dynamics
(Fig.~\ref{fig:decoding}c,d).

\begin{figure}[hp]
  \centering
  \includegraphics[width=\textwidth]{figs/decode_level.pdf}
  \caption{\textbf{Across-participant decoding accuracy varies with
      correlation order and cognitive engagement.}
    \textbf{a. Decoding accuracy as a function of order: PCA.} \textit{Order}
    ($x$-axis) refers to the maximum order of dynamic correlations that were
    available to the classifiers (see \textit{Feature weighting and
      testing}).  The reported across-participant decoding accuracies
    are averaged over all kernel shapes and widths (see
    \textit{Identifying robust decoding results}).  The $y$-values are
    displayed relative to chance accuracy.    The error bars
    denote 95\% confidence intervals across cross-validation folds
    (i.e., random assignments of participants to the training and test
    sets).  The colors denote the experimental condition.  Arrows
    denote sets of features that yielded reliably higher (upwards
    facing) or lower (downward facing) decoding accuracy than the mean
    of all other features (via a two-tailed test, thresholded at $p <
    0.05$).  The circled values represent the maximum decoding
    accuracy within each experimental condition.
    \textbf{b. Normalized decoding accuracy as a function of order: PCA.}
    This panel displays the same results as Panel a, but here each
    curve has been normalized to have a maximum value of 1 and and a
    minimum value of 0 (including the upper and lower bounds of the
    respective 95\% confidence intervals).  Panels a and b used PCA to
    project each high-dimensional pattern of dynamic correlations onto
    a lower-dimensional space.  \textbf{c.  Decoding accuracy as a
      function of order: eigenvector centrality.} This panel is in the
    same format as Panel a, but here eigenvector centrality has been
    used to project the high-dimensional patterns of dynamic
    correlations onto a lower-dimensional space.
    \textbf{d. Noramlized decoding accuracy as a function of order:
      eigenvector centrality.} This panel is in the same format as
    Panel b, but here eigenvector centrality has been
    used to project the high-dimensional patterns of dynamic
    correlations onto a lower-dimensional space.}
  \label{fig:decoding}
\end{figure}

@@
